Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-21 paper/paper.tex
Add Ewa's changest for section 6

In this paper we addressed the interesting and important new problem of 
scheduling and resource provisioning for scientific workflow ensembles 
on IaaS clouds. The problem is different from previous work on grid and 
utility grid scheduling in that cloud infrastructures provide more control 
over the resources, so the resource provisioning plan can be adjusted 
according to the requirements of the application. Therefore the problem 
space becomes larger: it requires not only an efficient mapping of tasks
to available resources, but also the selection of the best resource 
provisioning plan. Formulating the problem as a maximization of the
number of prioritized workflows completed from the ensemble also 
requires workflows to be admitted or rejected based on their estimated 
resource demands. We believe that this bi-constrained problem is highly 
relevant because such constraints are commonly imposed on many real-world 
projects.

We have considered both static and dynamic scheduling approaches and 
developed the SPDS, DPDS, WA-DPDS and SPSS algorithms based on these 
strategies. The algorithms were evaluated on ensembles of synthetic 
workflows, which were generated based on real scientific applications. For 
the purposes of evaluation we have developed a simulator that models: the 
cloud infrastructure, and a workflow engine with tightly-coupled scheduler 
and provisioner modules. The results of our simulation studies indicate 
that the two algorithms that take into account information about the 
structure of the workflow and task runtime estimates (WA-DPDS and SPSS) 
yield better results than the simple on-line priority-based scheduling 
strategy (DPDS).

During our work, several interesting lessons were learned about running 
workflow ensembles on cloud infrastructures. One result is that the simple 
strategy used by DPDS performs relatively well in terms of number of 
workflows completed and the high resource utilization it can achieve. However,
the extra workflows may be of lower-priority, and the high utilization
may be achieved by running tasks from workflows that cannot be completed
given the constraints. On the other hand, DPDS may be the only choice when 
estimates of task runtimes are not available in advance. We also observed
that the WA-DPDS and SPSS algorithms are able to find better quality 
solutions in terms of the target metrics: this is primarily a result of the 
admission algorithms rejecting workflows that are too large to be completed 
given the constraints, and accepting more smaller workflows that can be 
completed.

Another general observation is that although IaaS clouds provide elasticity
and the illusion of unlimited resources, which makes it possible to provision
more VMs to run workflows in parallel, it turns out that higher resource
utilization and better cost effectiveness can be achieved by limiting 
parallelism as much as possible. This observation comes from the fact that 
starting and stopping VMs incurs overheads which can be amortized by running
each VM for as long as possible. This effect was confirmed in the case of 
the SPSS algorithm for Montage and CyberShake workflows with small budgets
where SPSS opted to provision many VMs in parallel, and shut them down early,
so that tasks early in the ensemble workflows could be completed by their 
deadlines. Similarly, a high degree of parallelization leads to lower 
resource utilization and increased cost. This result, visible in 
Fig~\ref{fig:cost}, leads to multi-criteria decision problems and related 
trade-offs. E.g. our simulation results can be used as hints to answer
the question how much time and money is required to actually complete all the
workflows from the given ensemble.

Another lesson learned from our studies is the effect of task granularity on 
workflow ensembles. The benchmark workflows that we tested have relatively 
large number of tasks (50--1000), but the Montage and CyberShake workflows have 
fine-grained tasks with execution times on the order of seconds to minutes, 
while LIGO, Epigenomics and SIPHT have coarser-grained tasks with runtimes 
ranging from several minutes to hours. We observed that the dynamic 
algorithms achieved better results on the fine grained workflows since the 
simple random scheduling strategy leads to good load-balancing and high 
resource utilization for these ensembles. On the other hand, when tasks are
longer, it becomes more beneficial to apply the more sophisticated static 
scheduling strategy of SPSS.

Our current study suggests many areas for future work. One interesting
issue that we plan to investigate is how the static and dynamic algorithms 
perform with uncertain task runtimes and failures, which are known to have 
an important influence on overall system performance~\cite{Sakellariou2010,Dongarra2007}. 
We also plan to extend our application and infrastructure model to 
include the various data storage options available on clouds. A previous 
experimental study on that subject~\cite{Juve2010} suggests that the data 
demands of scientific workflows have a large impact on not only the 
execution time, but also on the cost of workflows in commercial clouds. 
Finally, we plan to investigate heterogeneous environments that include 
multiple VM types and cloud providers, including private and community clouds, 
which will make the problem even more complex and challenging~\cite{Marshall2010,vockler11,Juve2010}.

>>>>

In this paper we addressed the interesting and important new problem of 
scheduling and resource provisioning for scientific workflow ensembles 
on IaaS clouds. The problem is different from previous work on grid and 
utility grid scheduling in that cloud infrastructures provide more control 
over the resources, so the resource provisioning plan can be adjusted 
according to the requirements of the application. Therefore the problem 
space becomes larger: it requires not only an efficient mapping of tasks
to available resources, but also the selection of the best resource 
provisioning plan. Formulating the problem as a maximization of the
number of prioritized workflows completed from the ensemble is also novel and  
requires workflows to be admitted or rejected based on their estimated 
resource demands. We believe that this bi-constrained problem is highly 
relevant because such constraints are commonly imposed on many real-world 
projects. The approach is also directly applicable to grid environments 
that provide resource reservations and charge service units for resource use.

We have considered both static and dynamic scheduling approaches and 
developed the DPDS, WA-DPDS and SPSS algorithms based on these 
strategies. The algorithms were evaluated on ensembles of synthetic 
workflows, which were generated based on real scientific applications. For 
the purposes of evaluation we have developed a simulator that models: the 
cloud infrastructure, and a workflow engine with tightly-coupled scheduler 
and provisioner modules. The results of our simulation studies indicate 
that the two algorithms that take into account information about the 
structure of the workflow and task runtime estimates (WA-DPDS and SPSS) 
yield better results than the simple on-line priority-based scheduling 
strategy (DPDS).

During our work, several interesting lessons were learned about running 
workflow ensembles on cloud infrastructures. One result is that the simple 
strategy used by DPDS performs relatively well in terms of the number of 
workflows completed and the high resource utilization it can achieve. However,
the extra workflows may be of lower-priority, and the high utilization
may be achieved by running tasks from workflows that cannot be completed
given the constraints. On the other hand, DPDS may be the only choice when 
estimates of task runtimes are not available in advance. We also observed
that the WA-DPDS and SPSS algorithms are able to find better quality 
solutions in terms of the target metrics: this is primarily a result of the 
admission algorithms rejecting workflows that are too large to be completed 
given the constraints, and accepting more smaller workflows that can be 
completed.

Another general observation is that although IaaS clouds provide elasticity
and the illusion of unlimited resources, which makes it possible to provision
more VMs to run workflows in parallel, it turns out that higher resource
utilization and better cost effectiveness can be achieved by limiting 
parallelism. This observation comes from the fact that 
starting and stopping VMs incurs overheads which can be amortized by running
each VM for as long as possible. This effect was confirmed in the case of 
the SPSS algorithm for Montage and CyberShake workflows with small budgets.
In these cases SPSS decided to provision many VMs in parallel, and shut them down early,
so that tasks early in the ensemble workflows could be completed by their 
deadlines. Similarly, a high degree of parallelization leads to lower 
resource utilization and increased cost. This result, visible in 
Fig~\ref{fig:cost}, leads to multi-criteria decision problems and related 
trade-offs. E.g. our simulation results can be used as hints to answer
the question how much time and money is required to actually complete all the
workflows from the given ensemble.

Another lesson learned from our studies is the effect of task granularity on 
workflow ensembles. All the benchmark workflows that we tested have an average 
number of tasks (50--1000), but the Montage and CyberShake workflows have 
fine-grained tasks with execution times on the order of seconds to minutes, 
while LIGO, Epigenomics and SIPHT have coarser-grained tasks with runtimes 
ranging from several minutes to hours. We observed that the dynamic 
algorithms achieved better results on the fine grained workflows since the 
simple random scheduling strategy leads to good load-balancing and high 
resource utilization for these ensembles. On the other hand, when tasks are
longer, it becomes more beneficial to apply the more sophisticated static 
scheduling strategy of SPSS.

Our current study suggests many areas for future work. One interesting
issue that we plan to investigate is how the static and dynamic algorithms 
perform with uncertain task runtimes and failures, which are known to have 
an important influence on overall system performance~\cite{Sakellariou2010,Dongarra2007}. 
We also plan to extend our application and infrastructure model to explicitly 
include the various data storage options available on clouds (right now we only 
model data access as part of the task time and do not include data costs). A previous 
experimental study on that subject~\cite{Juve2010} suggests that the data 
demands of scientific workflows have a large impact on not only the 
execution time, but also on the cost of workflows in commercial clouds. 
Finally, we plan to investigate heterogeneous environments that include 
multiple VM types and cloud providers, including private and community clouds, 
which will make the problem even more complex and challenging~\cite{Marshall2010,vockler11,Juve2010}.
